### 2016-06-30 responseToReviewers.Rmd

9. In pg. 3 the authors propose that automated methods ``may allow for
identification of correlative patterns between WMH number, volume,
distribution, and disease state''. Some of these correlations have been
already identified in the literature, as the authors mention few lines
before. Moreover, the authors do not report number, volume, or
distribution of the lesions, to back up the utility of their method in
this regard.

> _\textcolor{blue}{We removed the sentence in question.  Also, as mentioned in
response to item 3, we added Table 1.}_

10. In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think this
is the right verb.

> _\textcolor{blue}{We agree that the sentence in question could be read this way.
We have, therefore, clarified the sentence in question to read
``Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
through each decision tree resulting in a single classification ``vote'' per tree.''}_

### 2016-06-30 responseToReviewers.tex

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{8}
\tightlist
\item
  In pg. 3 the authors propose that automated methods ``may allow for
  identification of correlative patterns between WMH number, volume,
  distribution, and disease state''. Some of these correlations have
  been already identified in the literature, as the authors mention few
  lines before. Moreover, the authors do not report number, volume, or
  distribution of the lesions, to back up the utility of their method in
  this regard.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We removed the sentence in question.  Also, as mentioned in
response to item 3, we added Table 1.}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{9}
\tightlist
\item
  In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think
  this is the right verb.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We agree that the sentence in question could be read this way.
We have, therefore, clarified the sentence in question to read
``Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
through each decision tree resulting in a single classification ``vote'' per tree.''}}
\end{quote}

### 2016-06-30 abstract.Rmd
ENH:  Submission files.

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with image features for
segmenting WMHs in a TBI cohort. The framework is built upon the
Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  Lesion load and overlap evaluative comparisons are
complimented by feature rankings which showcase the utility of the
concatenated approach.  Our findings suggest supervised learning methods
may be applied to quantify WMHs on routine brain imaging. Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

>>>>

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with image features for
segmenting WMHs in a TBI cohort. The framework is built upon the
Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  Performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.
Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
relative volume difference} = 43 $\pm$ 26\%.  In addition, three
lesion size ranges are selected to illustrate the variation in
performance with lesion size.  Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

### 2016-06-30 discussion.Rmd
ENH:  Submission files.

\textcolor{blue}{
In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies}
[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
[@vanginneken:2007aa].

>>>>

In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies
[@Debette:2010aa;@Kloppenborg:2014aa] exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference _between raters_ of 68 percent (cf Figure 5 (d))
[@vanginneken:2007aa].

### 2016-06-30 discussionX.Rmd
ENH:  Submission files.

In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies
[@Debette:2010aa;@Kloppenborg:2014aa] exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference _between raters_ of 68 percent (cf Figure 5 (d))
[@vanginneken:2007aa].

>>>>

\textcolor{blue}{
In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies}
[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference between raters of 68 percent (cf Figure 5 (d))}
[@vanginneken:2007aa].

### 2016-06-30 figureCaptions.Rmd
ENH:  Submission files.

__Figure 1:__ \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}

__Figure 2:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
the components relevant for this work).  Template construction is detailed in
[@Tustison:2015aa].  These images are important for asymmetry-based
features.

__Figure 3:__  Representation of Stage 1 feature images for subject 01C1019.  The
FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
The feature images are then generated for voxelwise input to the RF model which results in
the voting maps illustrated on the right.  This gives a probabilistic classification of
tissue type.  Not shown are the probability and voting images for the brain stem and
cerebellum.

@@

__Figure 5:__ \textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
volume difference.   These quantitative assessments are given for three quantile ranges
spanning the range of the manually-derived lesion volumes.    Overall improvement in all
three whole lesion-based measuers is seen as the
second Stage RF model is applied for all three quantile ranges.
The relative volume difference corresponding to the Stage 2 results tend to predict a
decreased predicted volume over the Stage 1 results.}

>>>>

__Figure 1:__ Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.

__Figure 2:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
from the MMRR data set (only shown are the FLAIR, T1, and T2 modalities---
the components relevant for this work).  These images are important for asymmetry-based
features.

__Figure 3:__  Representation of Stage 1 feature images for subject 01C1019.  The
FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
 to the space of the T1 image.  The three modality images are then preprocessed
(N4 bias correction and adaptive denoising) followed by
application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
the MMRR symmetric template and corresponding priors applied to the T1 image.
The feature images are then generated for voxelwise input to the RF model which results in
the voting maps illustrated on the right.  This gives a probabilistic classification of
tissue type.  Not shown are the probability and voting images for the brain stem and
cerebellum.

@@

__Figure 5:__ Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
volume difference.   These quantitative assessments are given for three quantile ranges
spanning the range of the manually-derived lesion volumes.    Overall improvement in all
three whole lesion-based measuers is seen as the
second Stage RF model is applied for all three quantile ranges.
The relative volume difference corresponding to the Stage 2 results tend to predict a
decreased predicted volume over the Stage 1 results.

### 2016-06-30 figureCaptions.tex
ENH:  Submission files.

\textbf{Figure 1:}
\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}

\textbf{Figure 2:} Canonical views of the mutlivariate, bilaterally
symmetric template constructed from the MMRR data set\textsuperscript{1}
(only shown are the FLAIR, T1, and T2 modalities--- the components
relevant for this work). Template construction is detailed
in\textsuperscript{2}. These images are important for asymmetry-based
features.

\textbf{Figure 3:} Representation of Stage 1 feature images for subject
01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
pre-aligned\textsuperscript{3} to the space of the T1 image. The three
modality images are then preprocessed (N4 bias
correction\textsuperscript{4} and adaptive denoising\textsuperscript{5})
followed by application of standard ANTs brain extraction and
\(n\)-tissue segmentation protocols using the MMRR symmetric template
and corresponding priors\textsuperscript{6} applied to the T1 image. The
feature images are then generated for voxelwise input to the RF model
which results in the voting maps illustrated on the right. This gives a
probabilistic classification of tissue type. Not shown are the
probability and voting images for the brain stem and cerebellum.

@@

\textbf{Figure 5:}
\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
volume difference.   These quantitative assessments are given for three quantile ranges
spanning the range of the manually-derived lesion volumes.    Overall improvement in all
three whole lesion-based measuers is seen as the
second Stage RF model is applied for all three quantile ranges.
The relative volume difference corresponding to the Stage 2 results tend to predict a
decreased predicted volume over the Stage 1 results.}

@@

\hypertarget{refs}{}
\hypertarget{ref-landman2011}{}
1. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.

\hypertarget{ref-Tustison:2015aa}{}
2. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
Grossman MC, Avants BB. Optimal symmetric multimodal templates and
concatenated random forests for supervised brain tumor segmentation
(simplified) with aNTsR. Neuroinformatics. 2015;13(2):209--25.

\hypertarget{ref-Avants:2014aa}{}
3. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
ToolKit image registration framework. Front Neuroinform. 2014;8:44.

\hypertarget{ref-Tustison:2010ac}{}
4. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
2010;29(6):1310--20.

\hypertarget{ref-Manjon:2010aa}{}
5. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
non-local means denoising of mR images with spatially varying noise
levels. J Magn Reson Imaging. 2010;31(1):192--203.

\hypertarget{ref-Tustison:2014ab}{}
6. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
and freeSurfer cortical thickness measurements. Neuroimage.
2014;99:166--79.

>>>>

\textbf{Figure 1:} Workflow illustration for the proposed pipeline.
Processing of the multi-modal input MRI for a single subject, using the
multi-modal symmetric template, results in the generation of the feature
images. These feature images are used as input to the Stage 1 RF model
producing the initial RF probability map estimates. The Stage 1 voting
maps, the original feature images, and the Stage 2 RF model result in
the final voting maps which includes the WMH probability estimate. Note
that the RF models are constructed once from a set of training data
which are processed using the same feature-construction pipeline as the
single-subject input MRI.

\textbf{Figure 2:} Canonical views of the mutlivariate, bilaterally
symmetric template constructed from the MMRR data set (only shown are
the FLAIR, T1, and T2 modalities--- the components relevant for this
work). These images are important for asymmetry-based features.

\textbf{Figure 3:} Representation of Stage 1 feature images for subject
01C1019. The FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
to the space of the T1 image. The three modality images are then
preprocessed (N4 bias correction and adaptive denoising) followed by
application of standard ANTs brain extraction and \(n\)-tissue
segmentation protocols using the MMRR symmetric template and
corresponding priors applied to the T1 image. The feature images are
then generated for voxelwise input to the RF model which results in the
voting maps illustrated on the right. This gives a probabilistic
classification of tissue type. Not shown are the probability and voting
images for the brain stem and cerebellum.

@@

\textbf{Figure 5:} Evaluation measures for both Stages of the
leave-one-out protocol of the described protocol in the Methods section:
(a) sensitivity, (b) positive predictive value, (c) \(F_1\) score, and
(d) relative volume difference. These quantitative assessments are given
for three quantile ranges spanning the range of the manually-derived
lesion volumes. Overall improvement in all three whole lesion-based
measuers is seen as the second Stage RF model is applied for all three
quantile ranges. The relative volume difference corresponding to the
Stage 2 results tend to predict a decreased predicted volume over the
Stage 1 results.

@@

\hypertarget{refs}{}

### 2016-06-30 format.Rmd
ENH:  Submission files.

header-includes:
   - \usepackage{booktabs}
   - \usepackage[final]{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{2}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
   - \definechangesauthor[name={James Stone}, color=magenta]{js}
   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
   - \definechangesauthor[name={David Tate}, color=orange]{dt}
bibliography: references.bib
csl: council-of-science-editors.csl
fontsize: 11pt
mainfont: Georgia
geometry: margin=1.0in
---

>>>>

header-includes:
   - \usepackage{booktabs}
   - \usepackage[final]{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{1.5}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
   - \definechangesauthor[name={James Stone}, color=magenta]{js}
   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
   - \definechangesauthor[name={David Tate}, color=orange]{dt}
bibliography: references.bib
csl: council-of-science-editors.csl
fontsize: 11pt
mainfont: Georgia
geometry: margin=1.0in
---

### 2016-06-30 intro.Rmd
ENH:  Submission files.

\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}

\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
classify ``outcome'' in TBI patients where ``outcome'' refers to the
spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
in patients who are comatose following severe TBI
the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].

\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.  Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical. The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.  In this work, we present an automated framework for quantification of WMHs in
multi-modal MRI using the random forest machine learning technique.}

@@

The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
normal human brain segmentation [@yi2009]) and other computer vision applications
(e.g., [@viola2005]).
\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
\textcolor{blue}{for both single and multi-modal acquisition protocols.}

@@

Random forests are conceptually straightforward [@breiman2001].
\textcolor{blue}{The basic component of
the random forest paradigm is the ``decision tree'' often represented by a flowchart or
graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.  The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.  For the proposed application, individual
voxels (and their corresponding feature values) are introduced at the root of a particular
decision tree and traverse the edges and internal nodes ultimately ending up at one of
the classification nodes according to the tests at each internal node.  A single random
forest model will consist of many such trees (often refered to as an ``ensemble'').}
\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
(e.g., AdaBoost)} [@schapire1990;@freund1997]
\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
\textcolor{blue}{Finally, Breiman} [@breiman2001]
\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
in the current random forest technique applied here.}
\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
each voxel location.}

>>>>

White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] populations.

Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa].
Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
classify "outcome" in TBI patients where "outcome" refers to the
spectrum of possible prognoses from death to disability to recovery [@Moen:2014aa]. Additionally,
in patients who are comatose following severe TBI
the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].

Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.  Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical. The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.  In this work, we present an automated framework for quantification of WMHs in
multi-modal MRI using the random forest machine learning technique.

@@

The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
normal human brain segmentation [@yi2009]) and other computer vision applications
(e.g., [@viola2005]).
Random forest-based paradigms have been successfully employed in the
delineation of other neuropathologies [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
for both single and multi-modal acquisition protocols.

@@

Random forests are conceptually straightforward [@breiman2001].
The basic component of
the random forest paradigm is the "decision tree" often represented by a flowchart or
graph where internal nodes represent "tests", or decisions, and the edges represent the outcome of those tests.  The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.  For the proposed application, individual
voxels (and their corresponding feature values) are introduced at the root of a particular
decision tree and traverse the edges and internal nodes ultimately ending up at one of
the classification nodes according to the tests at each internal node.  A single random
forest model will consist of many such trees (often refered to as an "ensemble").
Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
(e.g., AdaBoost) [@schapire1990;@freund1997]
influenced the similarly sytled conglomeration of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
Finally, Breiman [@breiman2001]
improved accuracy by random sampling of training data (i.e., "bagging") resulting
in the current random forest technique applied here.
As voxels and their feature values are "pushed" through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
each voxel location.
